### 2015-11-11 pde/pdetest1.tex
flip flop

u(x,t) = \sum_{n=1}^{\infty}e^{-k\left(\frac{n\pi}{L}\right)^{2}t}\sin{\left(\frac{n\pi x}{L}\right)}
& u(x,t) = \frac{a_{0}}{2} + \sum_{n=1}^{\infty}e^{-k\left(\frac{n\pi}{L}\right)^{2}t}\cos{\left(\frac{n\pi x}{L}\right)}\\

@@

u(x,t) =  \sum_{n=1}^{\infty}e^{-k\left(\left[n + \frac{1}{2}\right]\frac{\pi}{L}\right)^{2}t}\sin{\left(\left[n + \frac{1}{2}\right]\frac{\pi x}{L}\right)}
& u(x,t) =  \sum_{n=1}^{\infty}e^{-k\left(\left[n + \frac{1}{2}\right]\frac{\pi}{L}\right)^{2}t}\cos{\left(\left[n + \frac{1}{2}\right]\frac{\pi x}{L}\right)}

@@

\end{document}

>>>>

u(x,t) = \sum_{n=1}^{\infty}b_{n}e^{-k\left(\frac{n\pi}{L}\right)^{2}t}\sin{\left(\frac{n\pi x}{L}\right)}
& u(x,t) = \frac{a_{0}}{2} + \sum_{n=1}^{\infty}a_{n}e^{-k\left(\frac{n\pi}{L}\right)^{2}t}\cos{\left(\frac{n\pi x}{L}\right)}\\

@@

u(x,t) =  \sum_{n=1}^{\infty}a_{n}e^{-k\left(\left[n + \frac{1}{2}\right]\frac{\pi}{L}\right)^{2}t}\sin{\left(\left[n + \frac{1}{2}\right]\frac{\pi x}{L}\right)}
& u(x,t) =  \sum_{n=1}^{\infty}a_{n}e^{-k\left(\left[n + \frac{1}{2}\right]\frac{\pi}{L}\right)^{2}t}\cos{\left(\left[n + \frac{1}{2}\right]\frac{\pi x}{L}\right)}

@@

\newpage
\indent Now we will move on to the Cauchy wave equation. We will be presented with a similar problem: a wave equation on an unbounded domain. Consider the following problem:
\noindent \\\\
\noindent \textbf{\textit{Ex:}} Solve the following PDE
\[
\begin{cases}
u_{tt} - c^{2}u_{xx} = 0\\
u(x,0) = \sin{x}\\
u_{t}(x,0) = \cos{x}
\end{cases}
\]
\indent \textbf{\textit{Solution:}} Just as with the heat equation, we have a formal solution for the homogeneous wave equation. We will do the derivation, but I will tell you that the formal solution is
\[
u(x,t) = \frac{1}{2}\left[f(x+ct) + f(x-ct)\right] + \frac{1}{2c}\int_{x-ct}^{x+ct}g(s)\mathop{ds}.
\]
Now let's derive the formal solution. We want a solution in the form of
\[
u(x,t) = F(x+ct) + G(x-ct).
\]
Now we can say that $u_{t} = cF'(x+ct) - cG'(x-ct)$, and $u_{t}(x,0) = cF'(x) - cG'(x)$. This means
\begin{align*}
F(x) + G(x) &= f(x)\\
F(x) - G(x) &= \frac{1}{c}\int_{0}^{x}g(s)\mathop{ds} + c_{1}.
\end{align*}
Solving the system yields
\begin{align*}
u(x,t) &= \frac{1}{2} f(x+ct) + \frac{1}{2c}\int_{0}^{x+ct}g(s)\mathop{ds} + \frac{c_{1}}{2} + \frac{1}{2}f(x-ct) -\frac{1}{2c}\int_{0}^{x-ct}g(s)\mathop{ds} - \frac{c_{1}}{2}\\
&= \frac{1}{2}\left[f(x+ct) + f(x-ct)\right] + \frac{1}{2c}\int_{x-ct}^{x+ct}g(s)\mathop{ds},
\end{align*}
and that is the derivation. This problem is simply plugging into the formal solution. We have the following:
\begin{align*}
u(x,t) &= \frac{1}{2}\left[\sin{(x+ct)} + \sin{(x-ct)}\right] + \frac{1}{2c}\int_{x-ct}^{x+ct}\cos{(s)}\mathop{ds}\\
&= \frac{\sin{(x+ct)} + \sin{(x-ct)}}{2} + \frac{\cos{(x)}\sin{(ct)}}{c}\\
&=  \frac{\cos{(x)}\sin{(ct)}}{c} + \cos{(ct)}\sin{(x)},
\end{align*}
and we are done.

\newpage
\indent So far, we have only dealt with homogeneous Cauchy problems. We will now look at the case where we have a non-homogeneous problem. The first variation of this problem is the \textit{half-line} problem. In this problem, we will have a bound on our equation, and we will generalize the solution so that we are dealing with an unbounded Cauchy problem. Let's look at an example.
\noindent\\\\ \textbf{\textit{Ex:}} Solve the following PDE
\[
\begin{cases}
u_{t} - ku_{xx} = 0\\
u(x,0) = \phi(x)\\
u(0,t) = 0
\end{cases}
\]
\indent \textbf{\textit{Solution:}} We have both an initial condition and a boundary condition here. This problem is saying that at $x=0$, there is an impassable boundary. This forces our domain to be $[0,\infty)$. To solve a Cauchy problem, however, we want a domain of $(-\infty,\infty)$. We can force this domain by extending our initial condition, $\phi(x)$, over the boundary. We have the following rule:
\[
\phi_{0}(x) =
\begin{cases}
\phi(x)\quad\quad x>0\\
-\phi(-x)\ x<0.
\end{cases}
\]
Now we are dealing with the system
\[
\begin{cases}
u_{t} - ku_{xx} = 0\\
u(x,0) = \phi_{0}(x),
\end{cases}
\]
which we can solve using the formal solution to the heat equation. Recall that
\[
u(x,t) = \frac{1}{\sqrt{4k\pi t}}\int_{-\infty}^{\infty}e^{-\frac{(x-y)^{2}}{4kt}}\phi_{0}(y)\mathop{dy}.
\]
Substituting into the equation gives us
\begin{align*}
u(x,t) &= \frac{1}{\sqrt{4k\pi t}}\left[\int_{0}^{\infty}e^{-\frac{(x-y)^{2}}{4kt}}\phi(y)\mathop{dy} + \int_{-\infty}^{0}e^{-\frac{(x-y)^{2}}{4kt}}(-\phi(-y))\mathop{dy} \right]\\
&= \frac{1}{\sqrt{4k\pi t}}
\end{align*}
\end{document}
